---
title: "Native Land Attribution Quickie for Hack for LA"
author: Jude Calvillo
date: March 22, 2005
output: md_document
---

## Native Land Attribution Quickie for Hack for LA

```{r warning=F, message=F}
## Load libraries
library(tabulizer) # for extracting tables from PDFs
library(jsonlite) # for neatly converting JSON to dataframes
library(stringr) # for string splitting
library(dplyr) # for data transformations/aggregations
library(ggplot2) # for plotting
library(httr) # for raw API calls
library(ggmap) # easy interface to Google Geocode API
```

#### Parse Zip Code lookup table from LA County's official PDF  
Using Tabulizer library to extract tables from a PDF, then previewing the data.  
```{r}
## Parse the Zip code lookup table from LA County's PDF at: http://file.lacounty.gov/SDSInter/lac/1031552_MasterZipCodes.pdf
zips <- extract_tables("http://file.lacounty.gov/SDSInter/lac/1031552_MasterZipCodes.pdf")

## Preview
head(zips[[1]])
head(zips[[2]])
head(zips[[7]])
```

#### Clean up, normalize, and merge zip tables  
Header and column detection were a little problematic on a few pages, but the errors fall within a limited range of classes and the data is still usable. Therefore, let's just trim to what we need. This involves...    
+ Grabbing only the first two columns per table (given the commonality between the errors)  
+ Remove all rows that don't start with a zip code  
+ For all those rows with zip and area name merged, split the text and replace  
+ Flag areas that are officially part of the City of LA  
  
```{r}
## Generally, between the different types of errors we're seeing, the data we need is in the first 2 columns
## of the listed matrices, so let's row bind the first two columns of all the matrices
zips_df <- data.frame()
for(i in 1:length(zips)){
    zip_page <- as.data.frame(zips[i])[,c(1:2)]
    zips_df <- rbind(zips_df, zip_page)
}

## Pretty the names and remove rows that don't start with 5 numbers, then preview
names(zips_df) <- c("zip","area_name")
zips_df <- zips_df[grep("^[0-9][0-9][0-9][0-9][0-9].*", as.character(zips_df$zip)),]
head(zips_df)
tail(zips_df)

## How many rows have zip and area name merged?
zips_df$zip <- as.character(zips_df$zip)
zips_df$area_name <- as.character(zips_df$area_name)
length(zips_df$zip[grep(" [aA-zZ]", zips_df$zip)])

## For all those rows with zip and area name merged, split the text and replace
zips_df[grep(" [aA-zZ]", zips_df$zip),] <- str_split_fixed(zips_df$zip[grep(" [aA-zZ]", zips_df$zip)], " ", n = 2)
dim(zips_df)

## Flag whether an area is in the "City of LA" (vs in-line text), to clean and lean things out
zips_df$LA_city <- grepl("City of LA", zips_df$area_name, ignore.case = T)
zips_df$area_name <- gsub("\\(City of LA\\)| \\(City of LA\\)", "", zips_df$area_name, ignore.case = T)

## Preview final data
head(zips_df)
tail(zips_df)

```

#### Geocode (lat, lon) L.A. County zips using Google's Geocoding API  
Unfortunately, this API is rate-limited, so we'll factor that into our design.  
```{r}
if(!file.exists("dat/los-angeles_county_zip-codes_lat-lon.csv")){
  
    ## Grab my Google Geocode API key and register it with GGmaps, my easy interface to Google Maps family of APIs
    geo_key <- readRDS("geo_key.RDS")
    register_google(geo_key)
    
    ## Run zips through Geocode API call and keep lat/lon as new column values per zip
    zips_df[, c("lon", "lat")] <- geocode(zips_df$zip, output = "latlon")
    
    ## Save table for later use
    write.csv(zips_df, "dat/los-angeles_county_zip-codes_lat-lon.csv", row.names = F)
    
} else {
  
    ## Let user know we've already got the data and therefore don't need to call the API
    print("Zip to Lat/Lon lookup table already exists. Skipping API call, then previewing data...")
  
    ## Get existing data, if it's there
    zips_df <- read.csv("dat/los-angeles_county_zip-codes_lat-lon.csv", stringsAsFactors = F)
  
}

## Preview the data
head(zips_df)
tail(zips_df)

## Any NAs?
print("Possible problem zips, for inspection:")
zips_df[is.na(zips_df$lon),]


```

#### Send lat, lons to Native-land.ca API for list of Native lands LA County currently covers  
```{r}
## Test API call
test <- GET(paste0("https://native-land.ca/api/index.php?maps=languages,territories?positions=", zips_df$lat[1], ",", zips_df$lon[1]))
test            
```

