---
title: "Native Land Attribution Quickie for Hack for LA"
author: Jude Calvillo
date: April 29, 2021
output: md_document
---

## Native Land Attribution Quickie for Hack for LA

```{r warning=F, message=F}
## Load libraries
library(tabulizer) # for extracting tables from PDFs
library(jsonlite) # for neatly converting JSON to dataframes
library(stringr) # for string splitting
library(dplyr) # for data transformations/aggregations
library(ggplot2) # for plotting
library(httr) # for raw API calls
library(ggmap) # easy interface to Google Geocode API
```

### Parse Zip Code lookup table from LA County's official PDF  
Using Tabulizer library to extract tables from a PDF, then previewing the data.  
```{r}

## Parse the Zip code lookup table from LA County's PDF at: http://file.lacounty.gov/SDSInter/lac/1031552_MasterZipCodes.pdf
zips <- extract_tables("http://file.lacounty.gov/SDSInter/lac/1031552_MasterZipCodes.pdf")

## Preview
head(zips[[1]])
head(zips[[2]])
head(zips[[7]])
```

### Clean up, normalize, and merge zip tables  
Header and column detection were a little problematic on a few pages, but the errors fall within a limited range of classes and the data is still usable. Therefore, let's just trim to what we need. This involves...  

1. Grabbing only the first two columns per table (given the commonality between the errors)  
2. Remove all rows that don't start with a zip code  
3. For all those rows with zip and area name merged, split the text and replace  
4. Flag areas that are officially part of the City of LA  

--  
```{r}

## Generally, between the different types of errors we're seeing, the data we need is in the first 2 columns of the listed matrices, so let's row bind the first two columns of all the matrices
zips_df <- data.frame()
for(i in 1:length(zips)){
    zip_page <- as.data.frame(zips[i])[,c(1:2)]
    zips_df <- rbind(zips_df, zip_page)
}

## Pretty the names and remove rows that don't start with 5 numbers, then preview
names(zips_df) <- c("zip","area_name")
zips_df <- zips_df[grep("^[0-9][0-9][0-9][0-9][0-9].*", as.character(zips_df$zip)),]
head(zips_df)
tail(zips_df)

## How many rows have zip and area name merged?
zips_df$zip <- as.character(zips_df$zip)
zips_df$area_name <- as.character(zips_df$area_name)
length(zips_df$zip[grep(" [aA-zZ]", zips_df$zip)])

## For all those rows with zip and area name merged, split the text and replace
zips_df[grep(" [aA-zZ]", zips_df$zip),] <- str_split_fixed(zips_df$zip[grep(" [aA-zZ]", zips_df$zip)], " ", n = 2)
dim(zips_df)

## Flag whether an area is in the "City of LA" (vs in-line text), to clean and lean things out
zips_df$LA_city <- grepl("City of LA", zips_df$area_name, ignore.case = T)
zips_df$area_name <- gsub("\\(City of LA\\)| \\(City of LA\\)", "", zips_df$area_name, ignore.case = T)

## Preview final data
head(zips_df)
tail(zips_df)

```

#### Geocode (lat, lon) L.A. County zips using Google's Geocoding API  
Unfortunately, this API is rate-limited, so we'll factor that into our design.  
```{r}
if(!file.exists("dat/los-angeles-county_zip-codes_lat-lon.csv")){
  
    ## Grab my Google Geocode API key and register it with GGmaps, my easy interface to Google Maps family of APIs
    geo_key <- readRDS("geo_key.RDS")
    register_google(geo_key)
    
    ## Run zips through Geocode API call and keep lat/lon as new column values per zip
    zips_df[, c("lon", "lat")] <- geocode(zips_df$zip, output = "latlon")
    
    ## Save table for later use
    write.csv(zips_df, "dat/los-angeles_county_zip-codes_lat-lon.csv", row.names = F)
    
} else {
  
    ## Let user know we've already got the data and therefore don't need to call the API
    print("Zip to Lat/Lon lookup table already exists. Skipping API call, then previewing data...")
  
    ## Get existing data, if it's there
    zips_df <- read.csv("dat/los-angeles-county_zip-codes_lat-lon.csv", stringsAsFactors = F)
  
}

## Preview the data
head(zips_df)
tail(zips_df)

## Any NAs?
print("Possible problem zips, for inspection:")
zips_df[is.na(zips_df$lon),]


```

#### Send lat, lons to Native-land.ca API to get Native American lands and languages each area covers  
For each lat, lon combo, we ping Native-land.ca's API to get the Native American lands that once
```{r}
## Add empty territories and languages fields for now
zips_df$native_territories <- NA
zips_df$native_languages <- NA

## For each lat, lon combo, call the API
if(!file.exists("dat/los-angeles-county_native-american-lands.csv")){
  
    for(i in 1:length(zips_df$zip)){
  
        if(!is.na(zips_df$lon[i])){
          
          
          api_resp <- GET("https://native-land.ca/api/index.php", 
                          query = list(maps = "languages,territories", position = paste0(zips_df$lat[i], ",", zips_df$lon[i])),
                          add_headers('User-Agent' = 'Mozilla/5.0 (Windows NT 6.1; WOW64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/52.0.2743.116 Safari/537.36'))
          api_content <- fromJSON(rawToChar(api_resp$content))$properties
          zips_df$native_territories[i] <- paste0(api_content$Name[grep("territories", api_content$description)], collapse = ",")
          zips_df$native_languages[i] <- paste0(api_content$Name[grep("languages", api_content$description)], collapse = ",")
          
        } else {
          
          print(paste0("Not a valid set of coordinate values for area: ", zips_df$area_name[i], " (", zips_df$zip[i], ").", " Skipping."))
          
        }
    }
} else {
    
    ## Let user know we've already got the data and therefore don't need to call the API
    print("Table already exists. Skipping API call, then previewing data...")
  
    ## Get existing data, if it's there
    zips_df <- read.csv("dat/los-angeles-county_native-american-lands.csv", stringsAsFactors = F)
  
}


head(zips_df)

## Save to file
write.csv(zips_df, "dat/los-angeles-county_native-american-lands.csv", row.names = F)
```

### [Click to download lookup table of Los Angeles County Zip Codes to Native American Territories and Languages >>](https://raw.githubusercontent.com/judecalvillo/native-land-attribution/master/dat/los-angeles-county_native-american-lands.csv)




